Finiteness in Cubical Type Theory

### 2020-03-11 paper.tex
mostly addressed

Next, notice that the support lists of any two proofs of manifest Bishop
finiteness must be permutations of each other.
This will allow us to sort the support list of a proof of cardinal finiteness in
a coherently constant (definition~\ref{prop-trunc},
eliminator~\ref{elim-prop-coh}) way, pulling the support list out from the
truncation.
The cover proof emerges naturally from the definition of the permutation.
\subsection{Cardinal Finiteness and Kuratowski Finiteness}
\begin{romlemma} \label{cardinal-kuratowski}
  \begin{equation}
    \mathcal{C}(A) \simeq \mathcal{K}(A) \times \text{Discrete}(A)
  \end{equation}
\end{romlemma}
This proof is constructed by providing a pair of functions: one from
\(\mathcal{C}(A)\) to \(\mathcal{K}(A) \times \text{Discrete}(A)\), and one the
other way.
This pair implies an equivalence, because both source and target are
propositions.
The actual functions themselves are proven in our formalisation, as well as in
\cite{fruminFiniteSetsHomotopy2018}.
\section{Topos} \label{topos}
In this section we will prove that decidable Kuratowski finite types form a
\(\Pi\)-pretopos.
Along the way we will provide closure proofs for a number of the other
finiteness predicates.
As we saw in theorem~\ref{cardinal-kuratowski}, decidable Kuratowski finite
types are equivalent to cardinal finite types, so we will work with the latter
from now on.
Our first task is to show that cardinal finite types are closed under several
operations.
\subsection{Closure}
For the first three closure proofs, we only consider split enumerability:
as it is the strongest of the finiteness predicates, we can derive the other
closure proofs from it.

@@

From both of these types we can derive decidable equality.
\begin{romlemma}
  Any countable type has decidable equality.
\end{romlemma}
\subsection{Closure}
We know that countable infinity is not closed under the exponential (function
arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
what's left.
\begin{romtheorem} \label{split-countability-sigma}
  Split countability is closed under \(\Sigma\).
\end{romtheorem}
This proof does not mirror the proof of \(\Sigma\) closure on finite types,
surprisingly.
The pattern we used to pair up the relevant support lists for the finite types
would diverge for infinite types: instead, we need a \emph{pairing} function.

>>>>

Next, notice that the support lists of any two proofs of manifest Bishop
finiteness must be permutations of each other.
This will allow us to sort the support list of a proof of cardinal finiteness in
a coherently constant (definition~\ref{prop-trunc},
eliminator~\ref{elim-prop-coh}) way, pulling the support list out from the
truncation.
The cover proof emerges naturally from the definition of the permutation.
\subsection{Cardinal Finiteness and Kuratowski Finiteness}
\begin{romlemma} \label{cardinal-kuratowski}
  \begin{equation}
    \mathcal{C}(A) \simeq \mathcal{K}^f(A) \times \text{Discrete}(A)
  \end{equation}
\end{romlemma}
This proof is constructed by providing a pair of functions: one from
\(\mathcal{C}(A)\) to \(\mathcal{K}^f(A) \times \text{Discrete}(A)\), and one the
other way.
This pair implies an equivalence, because both source and target are
propositions.
The actual functions themselves are proven in our formalisation, as well as in
\cite{fruminFiniteSetsHomotopy2018}.
\section{Topos} \label{topos}
In this section we will prove that decidable Kuratowski finite types form a
\(\Pi\)-pretopos.
Along the way we will provide closure proofs for a number of the other
finiteness predicates.
As we saw in theorem~\ref{cardinal-kuratowski}, decidable Kuratowski finite
types are equivalent to cardinal finite types, so we will work with the latter
from now on.
Our first task is to show that cardinal finite types are closed under several
operations.
\subsection{Closure}
For the first three closure proofs, we only consider split enumerability:
as it is the strongest of the finiteness predicates, we can derive the other
closure proofs from it.

@@

From both of these types we can derive decidable equality.
\begin{romlemma}
  Any countable type has decidable equality.
\end{romlemma}
\subsection{Closure}
We know that countable infinity is not closed under the exponential (function
arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
what's left.
\begin{romtheorem} \label{split-countability-sigma}
  Split countability is closed under \(\Sigma\).
\end{romtheorem}
This proof does not mirror the proof of \(\Sigma\) closure on finite types.
The pattern we used to pair up the relevant support lists for the finite types
would diverge for infinite types: instead, we need a \emph{pairing} function.
